---
title: "DHIS2 Metadata Assessment"
author: "DHIS2 Implementation Team"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
always_allow_html: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(jsonlite, httr, purrr,knitr,magrittr,ggplot2,DT,dplyr,yaml,knitr,rmarkdown,dplyr,readr)


source("R/utils.R")
#Modify login details to the server as needed
baseurl <- Sys.getenv("baseurl")
username <- Sys.getenv("username")
password <- Sys.getenv("password")
include_slow <- ifelse(Sys.getenv("include_slow") == "TRUE",TRUE,FALSE)

views <- transformYAMLtoControlFile(include_protected = FALSE,include_slow = include_slow)

metadata <- transformYAMLtoMetadata(include_protected = FALSE,include_slow = include_slow)

#Login to the server
d2_session <- httr::handle(baseurl)
me <-httr::GET(paste0(baseurl,"api/me"), authenticate(username,password), handle= d2_session)


if (!me$status_code == 200L) {
  stop("Could not login to DHIS2")
}

getDHIS2Version <- function(d2_session = d2_session) {
    httr::GET(paste0(baseurl,"api/system/info"),  handle= d2_session) %>% 
    httr::content(.,"text") %>% 
    jsonlite::fromJSON(.) %>% 
    purrr::pluck("version")
}

checkVersion <- function(dhis2_version=getDHIS2Version()) {
  
  if (!grepl("2.36.4",dhis2_version)) {
    warning(paste("DHIS2 version not supported"))
  }
}

#checkVersion(getDHIS2Version(d2_session))

 r <- uploadMetadata(metadata, d2_session)
 if (r$status_code != 200L) {
   stop("Could not upload metadata!")
 }


 
```

# Target server: `r baseurl`


```{r sql_view_results, echo=FALSE,message=FALSE,results="asis"}

summary_views <- views$summary_uid

results <- purrr::map(views$summary_uid, function(x)  getSummaryView(x, d2_session = d2_session))
 
#Separate the results
#Errors 
errored_out <- unlist(lapply(results, function(x) all(class(x) == "list")))

if (any(errored_out)) {
results_with_errors <- results[errored_out] %>% 
  do.call("rbind.data.frame",. ) %>% 
  dplyr::left_join(views[,c("summary_uid","section","description")] , by =c("uid" = "summary_uid")) %>% 
  dplyr::select("Check" = description,
                "Message" = message)  
} else {
  results_with_errors <- data.frame(message="All checks ran successfu")
}

results_with_data <- do.call("rbind.data.frame",results[!errored_out])

results_table <- results_with_data %>% 
  dplyr::mutate(value = dplyr::coalesce(value,count,"0"),
                percent = sprintf(as.numeric(dplyr::coalesce(percent,"0")),fmt = '%#.2f')) %>% 
  dplyr::select(uid,value,percent) %>% 
  dplyr::left_join(views[,c("summary_uid","description","section","section_order","details_uid","severity")] , by =c("uid"="summary_uid")) %>% 
  dplyr::arrange(section,section_order) %>%
  dplyr::mutate(
    "Details" = paste0(
      "<a href='",
      d2_session$url,
      "api/sqlViews/",
      details_uid,
      "/data.csv' target='_blank'>Details</a>"
    ),
    severity = stringr::str_to_title(severity),
    percent = paste0(percent,"%")
  ) %>% 
  dplyr::select( "Theme" = section,
    "Issue" = description,
    "Level" = severity,
                "Count" = value,
                "Percentage" = percent,
    Details)
#Remove the views if requested to
if (Sys.getenv("cleanup_views") == TRUE) {
  r <- deleteMetadata(metadata, d2_session)
}
```


## Metadata issues 

```{r checks_with_results, echo=FALSE,message=FALSE}
# Metadata checks
DT::datatable(results_table,escape = FALSE)
```


## Checks which could not be run

```{r checks_with_errors, echo=FALSE,message=FALSE}
DT::datatable(results_with_errors, escape =  FALSE)
```


## Users

```{r users, echo=FALSE,message=FALSE}

url <- paste0(d2_session$url,"api/users?paging=false&fields=id,organisationUnits[level],userCredentials[userRoles[authorities],lastLogin,disabled]")
r <-httr::GET(url,handle=d2_session,timeout(600)) %>% 
  httr::content(.,"text") %>% 
  jsonlite::fromJSON()

n_users <-  length(r$users$id)

last_logins <- strptime(r$users$userCredentials$lastLogin,"%Y-%m-%dT%H:%M:%S") 

last_login_age <- last_logins %>% 
  lapply(., function(x) x[!is.na(x)]) %>% 
  lapply(., function(x) difftime(Sys.time(), x ,units="days"))


user_disabled <- r$users$userCredentials$disabled

user_levels <- r$users$organisationUnits

users_roles <- r$users$userCredentials$userRoles

user_has_super <- unlist(lapply(lapply(users_roles, function(x) stringr::str_detect(unlist(x),"^ALL$")),any))


```

- Total users: `r NROW(r$users$userCredentials)` 
- Total disabled users: `r sum(r$users$userCredentials$disabled,na.rm=TRUE)`
- Total users logged in within 7 days: `r sum(last_login_age <= 7,na.rm = TRUE)`
- Total users logged in within 30 days: `r sum(last_login_age <= 30,na.rm = TRUE)`
- Total users logged in within 90 days: `r sum(last_login_age <= 90,na.rm = TRUE)`
- Total users logged in within a year: `r sum(last_login_age <= 365,na.rm = TRUE)`
- Users who have never logged in: `r sum(is.na(last_logins))`
- Percentage of users who are disabled: `r round(sum(user_disabled)/length(user_disabled) * 100.0)` %
- Number of users who have superuser: `r sum(user_has_super)`
- Percentage of users who have superuser: `r round( sum(user_has_super)/ n_users * 100.0)` %

### Distribution of last user logins

```{r users_histogram,  echo=FALSE,message=FALSE, warning=FALSE}

qplot(as.numeric(last_login_age), 
geom="histogram",
main="User login age by day",
xlab="Days since last logged in",
ylab="Count")

```

### Distribution of users within the hierarchy

```{r users_hierarchy_level,  echo=FALSE,message=FALSE, warning=FALSE}

qplot(unlist(user_levels)[!user_disabled], 
geom="histogram",
main="Users hierarchy level",
xlab="Hierarchy level",
ylab="Count")

```


## Guidance section

This section of the report contains guidance information which provide more context
and expiation of each of the checks. 


```{r guidance, echo=FALSE,message=FALSE,results="asis"}

guidance <-  results_with_data %>% 
  dplyr::left_join(views[,c("summary_uid","description","section","section_order","details_uid","severity","introduction","recommendation")] , by =c("uid"="summary_uid")) %>% 
  dplyr::arrange(section,section_order) %>%
  dplyr::mutate(details = paste0("<a href='",
                                 d2_session$url,
                                 "api/sqlViews/",
                                 details_uid,
                                 "/data.csv' target='_blank'>Details</a>"),
                severity = stringr::str_to_title(severity)) 

sections <- unique(guidance$section)

for (i in seq_along(sections)) {
  this_section <- guidance %>% dplyr::filter(section == sections[i])
  cat(paste("##",sections[i]))
  cat("\n")
  cat("\n")
  for (j in seq_len(NROW(this_section))) {
    cat(paste0("### ",this_section$description[j]))
    cat("\n")
    cat("\n")
    cat(paste("Level:",this_section$severity[j]))
    cat("\n")
    cat("\n")
    cat(paste0("Issue count: ",this_section$value[j]," (",paste0(this_section$percent[j],"%"),")"))
    cat("\n")
    cat("\n")
    cat("**Introduction**")
    cat("\n")
    cat("\n")
    cat(this_section$introduction[j])
    cat("\n")
    cat("\n")
    cat("**Recommendation**")
    cat("\n")
    cat("\n")
    cat(this_section$recommendation[j])
    cat("\n")
    cat("\n")
    cat(this_section$details[j])
    cat("\n")
    cat("\n")
  }  
      
}

```



```{r native_checks, echo=FALSE,message=FALSE,results="asis"}

if (Sys.getenv("dhis2_checks") == "TRUE") {
  cat("\n")
  cat("\n")
  cat("## DHIS2 Integrity Checks")
  cat("\n")
  cat("\n")
   startNativeIntegrityCheck <- function(d2_session) {
   
   url <- paste0(d2_session$url,"api/dataIntegrity")
   r <- httr::POST(url,httr::content_type_json(),handle = d2_session) 
   
   if (r$status_code != 200L) {
     stop("Could not start DHIS2 integrity checks")
   }
   
   httr::content(r,"text") %>% 
   jsonlite::fromJSON(.) %>% 
   purrr::pluck("response") %>% 
   purrr::pluck("relativeNotifierEndpoint") %>% 
   stringr::str_replace(.,"^/","")
   
   }
   
   pollNativeIntegrityChecks <- function(endpoint,d2_session) {
   url <- paste0(d2_session$url,endpoint)
   r <- httr::GET(url,httr::content_type_json(),handle = d2_session)
  
  completed <- FALSE
  
  while(!completed) {
   Sys.sleep(5)
   
   completed <- httr::GET(url,httr::content_type_json(),handle = d2_session) %>% 
   httr::content("text") %>% 
   jsonlite::fromJSON(.) %>% 
   dplyr::pull(completed) %>% 
   any(.)
  
}
  
  #Now poll the system summary
  url <- paste0(d2_session$url,"api/system/taskSummaries/DATA_INTEGRITY/",substring(endpoint,nchar(endpoint)-10))
  
  httr::GET(url,httr::content_type_json(),handle=d2_session) %>% 
  httr::content("text") %>% 
  jsonlite::fromJSON(.)
   
 }
 
 endpoint <- startNativeIntegrityCheck(d2_session)

 r <- pollNativeIntegrityChecks(endpoint,d2_session) 

descriptions <- gsub('([[:upper:]])', ' \\1', names(r)) %>% 
stringr::str_to_sentence()
 
values <- lapply(r,length) %>%  unlist()

native_checks <- data.frame(Description = descriptions,Count = values,stringsAsFactors = FALSE)
row.names(native_checks) <- NULL

DT::datatable(native_checks,escape = FALSE)
}
 

```
